\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{geometry}
\usepackage{listings-solidity}  % Include Solidity highlighting
\usepackage{minted} % For advanced code listings

% Define a custom minted style (optional)
\usemintedstyle{colorful} % You can choose from various styles like 'monokai', 'tango', 'colorful', etc.

% Custom color setup
\definecolor{bashtextcolor}{RGB}{0, 0, 0} % Define black color

% Define a new command for inline code using minted
\newcommand{\codeinline}[1]{\mintinline{text}{#1}}

\geometry{a4paper, margin=1in}

\title{Smart Contracts Exercise 01: \\ Hello, Blockchain World!}
\author{}
\date{}

% Define a new command for inline code with a dark background
\newcommand{\codeblack}[1]{%
  \texttt{\colorbox{black!7}{\textcolor{black}{#1}}}%
}

% Define a new command for inline code with a dark background
\newcommand{\codegrey}[1]{%
  \texttt{\colorbox{black!4}{\textcolor{black}{#1}}}%
}

% Define custom colors (optional)
\definecolor{myURLColor}{RGB}{0, 102, 204} % Example: A shade of blue

\hypersetup{
    colorlinks=true,        % Enable colored links
    linkcolor=blue,         % Color for internal links (e.g., \ref, \cite)
    citecolor=blue,         % Color for citations
    filecolor=magenta,      % Color for file links
    urlcolor=myURLColor     % Color for external URLs
}

% Define a style for code listings
\lstdefinestyle{mystyle}{
    backgroundcolor=\color{lightgray!20},   
    commentstyle=\color{green!50!black},
    keywordstyle=\color{blue},
    numberstyle=\tiny\color{gray},
    stringstyle=\color{red},
    basicstyle=\ttfamily\footnotesize,
    breakatwhitespace=false,         
    breaklines=true,                 
    captionpos=b,                    
    keepspaces=true,                 
    numbers=left,                    
    numbersep=5pt,                  
    showspaces=false,                
    showstringspaces=false,
    showtabs=false,                  
    tabsize=2
}

\lstset{style=mystyle}
% Adding package for header and footer
\usepackage{fancyhdr}
\pagestyle{fancy}

% Define header and footer
\fancyhf{} % Clear current settings
\fancyhead[L]{Smart Contracts Exercise 01} % Left header
\fancyhead[R]{\thepage} % Right header with page number

\renewcommand{\headrulewidth}{0.4pt} % Line below header
% \renewcommand{\footrulewidth}{0.4pt} % Line above footer

\begin{document}

\maketitle
\noindent
Welcome to the first smart contracts exercise! In this exercise, you will become familiar with the basics of smart contract development. The goal is to create a simple smart contract. You will compile, test, and deploy this smart contract on a local network and, subsequently, deploy it to a live blockchain.

\section{Task: Set Up Hardhat Environment}

In this task, you will set up the Hardhat development environment. Hardhat is a
development environment for Ethereum software. It provides a suite of tools for
editing, compiling, debugging, and deploying your smart contracts. For this
exercise, you can choose between using a Docker container or installing locally
on your machine—select the option that best suits your development
preferences.

\subsection{Using Docker with VS Code}

This option uses Docker to create a development environment with all the
necessary tools and dependencies preinstalled.

\subsubsection*{Prerequisites:}
\begin{itemize}
    \item \textbf{\href{https://www.docker.com/products/docker-desktop}{Docker}} - A platform for developing, shipping, and running applications in containers.
    \item \textbf{\href{https://code.visualstudio.com/}{Visual Studio Code}} - A lightweight but powerful source code editor.
    \item \textbf{\href{https://marketplace.visualstudio.com/items?itemName=ms-vscode-remote.remote-containers}{Dev Containers}} - An extension to VS Code that lets you use a Docker container as a full-featured development environment.
\end{itemize}

\subsubsection*{Setting Up the Project:}

\begin{enumerate}
    \item Visit the following
          \href{https://github.com/radovluk/Smart-Contract-Exercise/tree/main/01-Hello-Blockchain-World/task/task-code}{GitHub repository} and clone it to your local machine.
    \item Open the repository folder in VS Code.
    \item When prompted, click ``Reopen in Container'' or use the command palette (F1)
          and run \codegrey{Dev Containers: Reopen in Container}.
\end{enumerate}

\noindent
Note: If you encounter permission issues when using Docker, you may need to adjust file permissions or run Docker with appropriate privileges. On Linux systems, you might need to add your user to the docker group: \codegrey{sudo usermod -aG docker \$USER} and then log out and log back in.

\medskip
\noindent
\textbf{Tip}: If you are using Windows, we strongly recommend using the Windows Subsystem for Linux (WSL) to follow this guide. For more information, refer to the \href{https://learn.microsoft.com/en-us/windows/wsl/about}{official documentation}.

\subsection{Local Setup}

If you prefer working directly on your machine without Docker, you can set up
the development environment locally. Before setting up Hardhat, ensure that you
have the following prerequisites installed on your system:

\subsubsection*{Prerequisites}
\begin{itemize}
    \item \textbf{Node.js}: \url{https://nodejs.org/en/} - An open-source, cross-platform, backend JavaScript runtime environment that runs on the V8 engine and executes JavaScript code outside a web browser.
    \item \textbf{NPM}: Node Package Manager, which comes with Node.js.
\end{itemize}

\noindent
Open your terminal and run the following commands to verify your installations:

\begin{minted}[bgcolor=gray!5, fontsize=\footnotesize]{bash}
$ node -v
$ npm -v
\end{minted}

\noindent
Both commands should return the installed version numbers of Node.js and NPM, respectively. Node.js provides the runtime environment required to execute JavaScript-based tools like Hardhat, while NPM is used to manage the packages and dependencies needed for development.

\medskip
\noindent
\textbf{Tip}: If you are using Visual Studio Code, consider installing the \href{https://marketplace.visualstudio.com/items?itemName=JuanBlanco.solidity}{Visual Studio Code Solidity Extension}. This extension helps your development process by providing features like syntax highlighting, code completion, etc.

\subsection{Creating a New Hardhat Project}

Create an empty working directory and then run the following commands to
initialize a Hardhat project:

\begin{minted}[bgcolor=gray!5, fontsize=\footnotesize]{bash}
$ npm init -y # Initialize an npm project in the directory.
$ npm install --save-dev hardhat # Install Hardhat in the directory.
$ npx hardhat init # Initialize a Hardhat project.
\end{minted}
Select the \codegrey{Create an empty hardhat.config.js} option with your keyboard
and press Enter.

\section{Task: Writing Your First Smart Contract}

Start by creating a new directory inside your project called \texttt{contracts}
and create a file inside this directory called \texttt{Greeter.sol}. Paste the
code below into the file and take a minute to read through it.

\noindent
\begin{minipage}[c]{\textwidth}
    \begin{lstlisting}[language=Solidity]
// File: contracts/Greeter.sol

// SPDX-License-Identifier: MIT
pragma solidity 0.8.28; // Specify the Solidity compiler version

/**
 * @title Greeter
 * @dev A simple smart contract that stores a greeting message.
 */
contract Greeter {
    string private greeting; // State variable to store the greeting message

    /**
    * @dev Constructor that sets the initial greeting message upon deployment.
    * @param _greeting The greeting message to be stored.
    */
    constructor(string memory _greeting) {
        greeting = _greeting;
    }

    /**
     * @dev Function to retrieve the greeting message.
     * @return The current greeting stored in the contract.
     */
    function greet() public view returns (string memory) {
        return greeting;
    }
}
\end{lstlisting}
\end{minipage}

The Greeter contract is a simple Solidity smart contract that stores a greeting
message, initializes it during deployment, and allows users to retrieve it via
a public function. To compile the contract, run \codegrey{npx hardhat compile}
in your terminal.

\begin{minted}[bgcolor=gray!5, fontsize=\footnotesize]{bash}
$ npx hardhat compile
Compiled 1 Solidity file successfully (evm target: paris).
\end{minted}

Hardhat compiles your Solidity smart contract and generates corresponding
artifacts—including the contract's ABI (Application Binary Interface, which
defines how to interact with the contract), bytecode (the compiled binary code
that runs on the Ethereum Virtual Machine), and related metadata—which are
stored in the \texttt{artifacts} folder. Take a look at the
\texttt{artifacts/contracts/Greeter.sol/Greeter.json} file.

\section{Task: Test your Smart Contract with Local \\ Hardhat Network}

\subsection{Set Up Hardhat-Toolbox Plugin}
In this task, you will write and execute a simple test case for the
\texttt{Greeter} contract using Hardhat's local network. For this task, you will
need the \texttt{@nomicfoundation/hardhat-toolbox} plugin, which integrates
testing libraries, Ethers.js, and other deployment utilities. Run the following
command in the project directory to install the plugin:

\begin{minted}[bgcolor=gray!5, fontsize=\footnotesize]{bash}
$ npm install --save-dev @nomicfoundation/hardhat-toolbox
\end{minted}

\noindent
To include the plugin in your Hardhat project, modify your \texttt{hardhat.config.js} file in the project directory so that it looks like this:

\begin{minted}[bgcolor=gray!5, fontsize=\footnotesize]{javascript}
// File: hardhat.config.js

require("@nomicfoundation/hardhat-toolbox");

/** @type import('hardhat/config').HardhatUserConfig */
module.exports = {
  solidity: "0.8.28",
};
\end{minted}
For more information about plugins and how to test contracts in Hardhat, visit the
\\ \href{https://hardhat.org/tutorial/testing-contracts}{Hardhat
    documentation}.

\subsection{Writing a Simple Test}

Create a new directory named \texttt{test} in your project root and add a file
called \texttt{Greeter.js} with the following content:

\noindent
\begin{minipage}[c]{\textwidth}
    \begin{minted}[bgcolor=gray!5, fontsize=\footnotesize]{javascript}
// File: test/Greeter.js

// Import the 'expect' function from Chai for assertions
const { expect } = require("chai");

// Test suite for the Greeter contract
describe("Greeter contract says Hello, Blockchain World!", function () {
  
  // Test to ensure the initial greeting is set correctly upon deployment
  it("Should set the initial greeting correctly.", async function () {
    
    // Define the initial greeting message
    const initialGreeting = "Hello, Blockchain World!";
    
    // Deploy the Greeter contract with the initial greeting
    const greeter = await ethers.deployContract("Greeter", [initialGreeting]);
    
    // Wait for the deployment to complete
    await greeter.waitForDeployment();

    // Retrieve the stored greeting from the contract
    const greeting = await greeter.greet();

    // Verify that the retrieved greeting matches the initial greeting
    expect(greeting).to.equal(initialGreeting);
  });
});
\end{minted}
\end{minipage}

\noindent
\textbf{Note:} For all exercises, we will use \href{https://docs.ethers.org/v6/}{ethers.js v6}. If you're using an older version of ethers, the syntax may differ. For example, in ethers v5, you would use \\ \codegrey{await ethers.getContractFactory("Greeter")} and then \\ \codegrey{await greeterFactory.deploy(initialGreeting)} instead. Keep this in mind if you encounter errors.

\subsection{Running the Test}

Execute the test on Hardhat's local network by running the following commands
in your terminal:

\begin{minted}[bgcolor=gray!5, fontsize=\footnotesize]{bash}
$ npx hardhat test
\end{minted}
Congratulations! You wrote, compiled, and tested your first smart contract!

\section{Task: Deploying to a Live Network}

Once you have programmed and tested your dApp, you will want to deploy it to a
public blockchain so others can access it. For the purposes of our
exercise, we will not use the Ethereum mainnet because it would require paying
with real money, but instead we will use a live testnet. A testnet mimics real-world
scenarios without risking actual funds. Ethereum has several
\href{https://ethereum.org/en/developers/docs/networks/#ethereum-testnets}{testnets};
for our purposes, we will use the \href{https://sepolia.dev/}{Sepolia testnet}. Deploying to a testnet is the same as deploying to the mainnet at the
software level. The only difference is the network you connect to.

\subsection{Prerequisites}

To complete this task, you will need the following tools:

\begin{itemize}
    \item \textbf{MetaMask}: A popular Ethereum wallet that allows you to interact with the Ethereum blockchain. You can download the MetaMask extension for your browser from the \href{https://metamask.io/}{official website} and set it up. Alternatively, you can use other Ethereum wallets or simply create your own private-public key pair.

    \item \textbf{Infura API Key}: Infura provides access to Ethereum nodes without the need to run your own. Sign up at \href{https://infura.io/}{Infura} to obtain an API key.

    \item \textbf{Sepolia Faucet}: Acquire Sepolia test Ether from a faucet to fund your deployment. Even on testnets, you'll need testnet ETH to pay for gas fees. Make sure you have enough Sepolia ETH (0.01 Sepolia ETH should be sufficient for this exercise) in your wallet before deployment. Gas prices fluctuate based on network congestion, even on testnets. Some reliable faucets include:
          \begin{itemize}
              \item \href{https://cloud.google.com/application/web3/faucet/ethereum/sepolia}{\textbf{Google Cloud Web3}} (needs only Google account)
              \item \href{https://docs.metamask.io/developer-tools/faucet/}{Metamask Sepolia Faucet} (needs some ETH on mainnet)
              \item \href{https://www.alchemy.com/faucets/ethereum-sepolia}{Alchemy Sepolia Faucet}  (needs some ETH on mainnet)
          \end{itemize}
\end{itemize}

\subsection{Configuring Hardhat for Sepolia Deployment}

To deploy your smart contract to the Sepolia testnet, you need to configure
Hardhat with the network details and your wallet credentials.

\subsubsection*{Storing Sensitive Information}

It's crucial to keep sensitive information like your private key and Infura API
key secure. We recommend using configuration variables to manage these
credentials, but only for the purpose of this exercise. A Hardhat project can use
configuration variables for user-specific values or for data that shouldn't be
included in the code repository. These variables are set via tasks in the vars
scope and can be retrieved in the config using the vars object.

\begin{itemize}
    \item Set the INFURA\_API\_KEY

          \begin{minted}[bgcolor=gray!5, fontsize=\footnotesize]{bash}
    $ npx hardhat vars set INFURA_API_KEY
    Enter value: ********************************
\end{minted}

    \item Set the SEPOLIA\_PRIVATE\_KEY

          \begin{minted}[bgcolor=gray!5, fontsize=\footnotesize]{bash}
    $ npx hardhat vars set SEPOLIA_PRIVATE_KEY
    Enter value: ********************************
\end{minted}

\end{itemize}

\begin{itemize}
    \item \textbf{Warning}: Configuration variables are stored in plain text on your disk. Avoid using this feature for data you wouldn't normally save in an unencrypted file. Run \texttt{npx hardhat vars path} to find the storage file location. Never use your private key associated with real money in plain text!
\end{itemize}

\subsubsection*{Updating \texttt{hardhat.config.js}}

Modify your \texttt{hardhat.config.js} file to include the Sepolia network
configuration:

\noindent
\begin{minipage}[c]{\textwidth}
    \begin{minted}[bgcolor=gray!5, fontsize=\footnotesize]{javascript}
// File: hardhat.config.js

require("@nomicfoundation/hardhat-toolbox");

const INFURA_API_KEY = vars.get("INFURA_API_KEY");
const SEPOLIA_PRIVATE_KEY = vars.get("SEPOLIA_PRIVATE_KEY");

/** @type import('hardhat/config').HardhatUserConfig */
module.exports = {
  solidity: "0.8.28",
  networks: {
    sepolia: {
      url: `https://sepolia.infura.io/v3/${INFURA_API_KEY}`,
      accounts: [`0x${SEPOLIA_PRIVATE_KEY}`],
    },
  },
};
\end{minted}
\end{minipage}

\noindent
This configuration tells Hardhat how to connect to the Sepolia testnet using your Infura API key and how to deploy contracts using your wallet's private key.

\subsection{Deploying the Smart Contract to Sepolia}

With the configuration in place, you're ready to deploy your smart contract to
the Sepolia testnet.

\subsubsection{Creating a Deployment Script}

\begin{itemize}
    \item 1. Create a \texttt{scripts} Directory: In your project root, create a new directory named \texttt{scripts}.

    \item 2. Add a Deployment Script: Inside the \texttt{scripts} directory, create a file named \texttt{deploy.js} and add the following content:
\end{itemize}

\noindent
\begin{minipage}[c]{\textwidth}
    \begin{minted}[bgcolor=gray!5, fontsize=\footnotesize]{javascript}
// File: scripts/deploy.js

const hre = require("hardhat");

async function main() {
    // Set the initial greeting message
    const initialGreeting = "Hello, Blockchain World!";
    
    // Deploy the Greeter contract with the initial greeting
    const greeter = await ethers.deployContract("Greeter", [initialGreeting]);
    console.log(`Greeter contract deployed to: ${greeter.target}`);

    // Wait for the deployment to complete
    await greeter.waitForDeployment();

    // Retrieve the stored greeting from the contract
    const greeting = await greeter.greet();
    console.log(`Contract greeting: ${greeting}`);
}

main()
    .then(() => process.exit(0))
    .catch((error) => {
        console.error(error);
        process.exit(1);
    });
\end{minted}
\end{minipage}

\subsubsection{Executing the Deployment}

Run the deployment script using Hardhat with the Sepolia network specified:

\begin{minted}[bgcolor=gray!5, fontsize=\footnotesize]{bash}
$ npx hardhat run scripts/deploy.js --network sepolia
\end{minted}
Expected Output:
\begin{minted}[bgcolor=gray!5, fontsize=\footnotesize]{bash}
Greeter contract deployed to: <ContractAddress>
Contract greeting: Hello, Blockchain World!
\end{minted}
You can verify the deployment by visiting the Sepolia Etherscan explorer and
searching for your contract address:
\url{https://sepolia.etherscan.io/address/<ContractAddress>}. Also, search for
your account address to see your interactions with the deployed contract.

\noindent
Hardhat also includes Hardhat Network, a local Ethereum network node for development. It enables you to deploy contracts, run tests, and debug code, all within your local environment. We already used it when running our test. To use it explicitly, open a separate terminal and run \codegrey{npx hardhat node}. To deploy the contract, run \codegrey{npx hardhat run scripts/deploy.js --network hardhat} in another terminal. See the \href{https://hardhat.org/hardhat-network/docs/overview\#hardhat-network}{Hardhat Network} for more information.

\subsection{Interacting with Your Deployed Contract}

Now that your contract is live on the Sepolia testnet, you can interact with it
using various tools:

\begin{itemize}
    \item \textbf{Etherscan}: View contract details, read functions, and execute transactions directly from the Etherscan interface.

    \item \textbf{Web3 Interfaces}: Integrate your contract with frontend applications using libraries like \texttt{ethers.js} or \texttt{web3.js}.

    \item \textbf{Hardhat Tasks}: Write scripts or use the \href{https://hardhat.org/hardhat-runner/docs/guides/hardhat-console}{Hardhat console} to interact programmatically with your contract.
\end{itemize}
Tip: If you run the deployment script without specifying the \texttt{--network} parameter, it will deploy to the local Hardhat network.
\begin{minted}[bgcolor=gray!5, fontsize=\footnotesize]{bash}
$ npx hardhat run scripts/deploy.js
\end{minted}

\subsection{Further Reading}

For more detailed information, refer to the following resources:

\begin{itemize}
    \item \href{https://docs.soliditylang.org/en/latest/}{Solidity Documentation}
    \item \href{https://hardhat.org/docs}{Hardhat Documentation}
    \item \href{https://solidity-by-example.org/}{Solidity by Example}
    \item \href{https://docs.ethers.org/v6/}{Ethers.js Documentation} (for scripting)
    \item \href{https://www.chaijs.com/}{Chai Assertion Library}

\end{itemize}
Congratulations! You have successfully deployed your first smart contract to a live blockchain network! Stay tuned for the upcoming exercises!

\end{document}
